---
title: "Penguins"
author: "Kasia"
date: "2023-05-31"
output:
  pdf_document: default
  html_document: default
---

### Loading packages

```{r loading data}
library(tidyverse)
library(palmerpenguins)
View(penguins)

```
#### Inspecting data
```{r getting to know the data}
head(penguins)
str(penguins)
colnames(penguins)
```
#### Summarising data without NA values
```{r removing NA values}
penguins %>%
  group_by(species, island) %>%
  drop_na() %>%
  summarise(max_bl = max(bill_length_mm), mean_bl = mean(bill_length_mm))
```
#### Transforming data - adding a column with kilograms
```{r adding a column with kilograms instead of grams}
penguins %>%
  mutate(body_mass_kg = body_mass_g/100)
```
#### Visualising data with ggplot2 - showing different penguins species
```{r loading ggplot}
library(ggplot2)
library(palmerpenguins)
```


```{r visualising data}
data("penguins")
ggplot(data = penguins) +
  geom_point(mapping = aes(x = flipper_length_mm, y = body_mass_g))
```
#### Mapping a new variable to a new aesthetic
```{r}
ggplot(data = penguins) +
  geom_point(mapping = aes(x = flipper_length_mm, y = body_mass_g, shape = species, color = species))
```
#### Showing the relationship between the trend line and data point more clearly
```{r showing relationship between variables}
ggplot(data = penguins) +
  geom_smooth(mapping = aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(mapping = aes(x = flipper_length_mm, y = body_mass_g))
```
#### Plotting a separate line for each species of penguins
```{r separate line for each species}
ggplot(data = penguins) +
  geom_smooth(mapping = aes(x = flipper_length_mm, y = body_mass_g, linetype = species))
```

# Jittering a plot
```{r adding a random noise}
ggplot(data = penguins) +
  geom_jitter(mapping = aes(x = flipper_length_mm, y = body_mass_g))
```
# Focusing on the data for each species of penguins
```{r showing data for each species}
ggplot(data = penguins) +
  geom_point(mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  facet_grid(~species)
```
####  Labeling the plot
```{r naming the plot}
ggplot(data = penguins) +
  geom_point(mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  labs(title = "Palmer Penguins: Body Mass vs Flipper Length",
       subtitle = "Sample of Three Penguin Species",
       caption = "Data collected by Dr.Kristen Gorman") +
  annotate("text", x = 220, y = 3500, label = "The Gentoos are the largest")

```

`````{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

